---
title: "LSOAs"
author: "Nads"
date: "07/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(dplyr)
library(readr)
library(janitor)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(readxl)
library(tsibble)
library(forecast)
library(sf)
library(ggspatial)
library(ggmap)
library(tmap)


## lOAD DATASETS  ------------------------------------------------------------------- 
calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>% clean_names()
missing <- subset(calls, incident_type == "Missing Person")
missing <- subset(missing, select = -c(4, 6))
 
  #add lsoa 
name_code <- read_csv("data/Lower_Layer_Super_Output_Areas_December_2011_Names_and_Codes_in_England_and_Wales.csv") %>%
  clean_names() %>%
  select(lsoa11cd, lsoa11nm)
missing <- left_join(missing, name_code, by = c("lsoa" = "lsoa11nm"))


```





## Map Missper across LSOAS

```{r}

###
#lsoa shapefile 
shp <- file.choose()  #read boundary shapefile
england_lsoa <- st_read(shp)
ggplot() + 
  geom_sf(data = england_lsoa)
england_lsoa <- st_transform(england_lsoa, 4326)

calls_per_lsoa <- missing %>%  #per name not code
  group_by(lsoa11cd) %>%
  summarise(count=n())

cheshire_lsoa <- left_join(england_lsoa, calls_per_lsoa, by = c("code"="lsoa11cd"))

ggplot() + 
  annotation_map_tile() +  # add basemap
  geom_sf(data = cheshire_lsoa, aes(fill = count), alpha = 0.7) + # alpha sets the opacity
  scale_fill_gradient2(name ="Number of Missper") + 
  labs(
    title = "Count of MissPer by LSOA",
    x = NULL,
    y = "Proportion of Missing Person",
    fill = "Missing Person"
  )


## tmap 

current_style <- tmap_style("col_blind")

tm_shape(cheshire_lsoa) + 
  tm_fill("count", title = "Missper") +
  tm_borders(alpha = 0.1) +
  tm_layout(main.title = "MissPer: 2015:2020", main.title.size = 0.7 ,
            legend.position = c("right", "bottom"), legend.title.size = 0.8)



## jenks 
tm_shape(cheshire_lsoa) + 
  tm_fill("count", style = "jenks", title = "Missper") +
  tm_borders(alpha = 0.1) +
  tm_layout(main.title = "MissPer: 2015:2020", main.title.size = 0.7 ,
            legend.position = c("right", "bottom"), legend.title.size = 0.8)



```

There is a siginificant distribution of counts of missper across cheshire




## LISA 


```{r}
library(spdep)

##global 

# generate weight list using Queen 
missing_m <- as(cheshire_lsoa, "Spatial")
w <- poly2nb(missing_m, row.names=missing_m$code)
ww <-  nb2listw(w, style='W')

moran.plot(missing_m$count, ww) 
#positive spatial autocorrelation 

moran(missing_m$count, ww, n=length(ww$neighbours), S0=Szero(ww))
#0.2462985

moran.mc(missing_m$count, ww, nsim=99999)


## local 

locm_bm <- localmoran(missing_m$count, ww)
summary(locm_bm)

missing_m$s_missing <- scale(missing_m$count) %>% as.vector() #scaled
missing_m$l_missing <- lag.listw(ww, missing_m$s_missing)  #lagged

x <- missing_m$s_missing  #check scatter plot 
y <- missing_m$l_missing
xx <- tibble(x,y)
moran.plot(x, ww)

missing_m <- st_as_sf(missing_m) %>% 
  mutate(quad_sig = ifelse(missing_m$s_missing > 0 & 
                              missing_m$l_missing > 0 & 
                              locm_bm[,5] <= 0.05, 
                     "high-high",
                     ifelse(missing_m$s_missing <= 0 & 
                              missing_m$l_missing <= 0 & 
                              locm_bm[,5] <= 0.05, 
                     "low-low", 
                     ifelse(missing_m$s_missing > 0 & 
                              missing_m$l_missing <= 0 & 
                              locm_bm[,5] <= 0.05, 
                     "high-low",
                     ifelse(missing_m$s_missing <= 0 & 
                              missing_m$l_missing > 0 & 
                              locm_bm[,5] <= 0.05,
                     "low-high", 
                     "non-significant")))))

table(missing_m$quad_sig)  #54 high-high, 598 non-significant 
nrow(locm_bm[locm_bm[,5] <= 0.05,]) #only 54 are less then 0.05

qtm(missing_m, fill="quad_sig", fill.title="LISA")



```



Using the queens criteria, created a weighted list matrix

There is a positive spatial autocorrelation. with a moran value of 0.23 and a p value of of 1e-05 we can conclude that there is indeed significant global spatial autocorrelation (the spatial processes promoting the observed pattern of values is random by chance), so what is driving this measure. 


LISA map is telling us that there was some moderate statistically significant clustering of mp in Cheshire


This paper will test to see if levels of vulnerability can explain these levels using 
  - IMD
  - MH
  - maybe carehomes 
  
  
  
  
  
  
  
  
  
  



